---
title: "Introduction to FedData"
output: 
  rmarkdown::html_document:
    theme: simplex
    toc: true
    toc_depth: 4
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
```

`FedData` is a package for downloading data from federated (usually US Federal) data sets, using a consistent, spatial-first API. In this vignette, you'll learn how to load the `FedData` package, download data for an "template" area of interest, and make simple web maps using the [`mapview`](https://r-spatial.github.io/mapview/) package.

## Load `FedData` and define a study area
```{r setup, message=FALSE}
# Load FedData and magrittr
library(FedData)
library(magrittr)
library(terra)

# Install mapview if necessary
if (!require("mapview")) {
  install.packages("mapview")
}

library(mapview)
mapviewOptions(
  basemaps = c(),
  homebutton = FALSE,
  query.position = "topright",
  query.digits = 2,
  query.prefix = "",
  legend.pos = "bottomright",
  platform = "leaflet",
  fgb = TRUE,
  georaster = TRUE
)

# Create a nice mapview template
plot_map <-
  function(x, ...) {
    if (inherits(x, "SpatRaster")) {
      x %<>%
        as("Raster")
    }

    bounds <-
      FedData::meve %>%
      sf::st_bbox() %>%
      as.list()

    mapview::mapview(x, ...)@map %>%
      leaflet::removeLayersControl() %>%
      leaflet::addTiles(
        urlTemplate = "https://basemap.nationalmap.gov/ArcGIS/rest/services/USGSShadedReliefOnly/MapServer/tile/{z}/{y}/{x}"
      ) %>%
      leaflet::addTiles(
        urlTemplate = "https://tiles.stadiamaps.com/tiles/stamen_toner_lines/{z}/{x}/{y}.png"
      ) %>%
      leaflet::addTiles(
        urlTemplate = "https://tiles.stadiamaps.com/tiles/stamen_toner_labels/{z}/{x}/{y}.png"
      ) %>%
      # leaflet::addProviderTiles("Stamen.TonerLines") %>%
      # leaflet::addProviderTiles("Stamen.TonerLabels") %>%
      leaflet::addPolygons(
        data = FedData::meve,
        color = "black",
        fill = FALSE,
        options = list(pointerEvents = "none"),
        highlightOptions = list(sendToBack = TRUE)
      ) %>%
      leaflet::fitBounds(
        lng1 = bounds$xmin,
        lng2 = bounds$xmax,
        lat1 = bounds$ymin,
        lat2 = bounds$ymax
      )
  }


# FedData comes loaded with the boundary of Mesa Verde National Park, for testing
FedData::meve

plot_map(FedData::meve,
  legend = FALSE
)
```

## Datasets

### USGS National Elevation Dataset
```{r NED, message=FALSE}
# Get the NED (USA ONLY)
# Returns a raster
NED <- get_ned(
  template = FedData::meve,
  label = "meve"
)

plot_map(NED,
  layer.name = "NED Elevation (m)",
  maxpixels = terra::ncell(NED)
)
```

### ORNL Daymet
```{r DAYMET, message=FALSE, warning=FALSE}
# Get the DAYMET (North America only)
# Returns a raster
DAYMET <- get_daymet(
  template = FedData::meve,
  label = "meve",
  elements = c("prcp", "tmax"),
  years = 1985
)

plot_map(DAYMET$tmax$`1985-10-23`,
  layer.name = "TMAX on 23 Oct 1985 (Degrees C)",
  maxpixels = terra::ncell(DAYMET$tmax)
)
```

### NOAA GHCN-daily
```{r GHCN-precipitation, message=FALSE}
# Get the daily GHCN data (GLOBAL)
# Returns a list: the first element is the spatial locations of stations,
# and the second is a list of the stations and their daily data
GHCN.prcp <- get_ghcn_daily(
  template = FedData::meve,
  label = "meve",
  elements = c("prcp")
)

GHCN.prcp$spatial %>%
  dplyr::mutate(label = paste0(ID, ": ", NAME)) %>%
  plot_map(
    label = "label",
    layer.name = "GHCN_prcp",
    legend = FALSE
  )
```

#### Standardized data
```{r GHCN-temperature, message=FALSE}
# Elements for which you require the same data
# (i.e., minimum and maximum temperature for the same days)
# can be standardized using standardize==T
GHCN.temp <- get_ghcn_daily(
  template = FedData::meve,
  label = "meve",
  elements = c("tmin", "tmax"),
  years = 1980:1985,
  standardize = TRUE
)

GHCN.temp$spatial %>%
  dplyr::mutate(label = paste0(ID, ": ", NAME)) %>%
  plot_map(
    label = "label",
    layer.name = "GHCN_temp",
    legend = FALSE
  )
```

### USGS National Hydrography Dataset
```{r NHD, message=FALSE}
# Get the NHD (USA ONLY)
NHD <-
  get_nhd(
    template = FedData::meve,
    label = "meve",
    force.redo = TRUE
  )

# FedData comes with a simple plotting function for NHD data
plot_nhd(NHD, template = FedData::meve)

# Or, do an interactive map
NHD[purrr::map_lgl(NHD, ~ (nrow(.x) != 0))] %>%
  plot_map()

# You can also retrieve the Watershed Boundary Dataset
WBD <-
  get_wbd(
    template = FedData::meve,
    label = "meve"
  )

plot_map(WBD,
  zcol = "name",
  legend = FALSE
)
```

### USDA NRCS SSURGO
```{r SSURGO, message=FALSE, warning=FALSE}
# Get the NRCS SSURGO data (USA ONLY)
SSURGO.MEVE <- get_ssurgo(
  template = FedData::meve,
  label = "meve"
)

SSURGO.MEVE$spatial %>%
  rmapshaper::ms_simplify() %>%
  dplyr::left_join(
    SSURGO.MEVE$tabular$mapunit %>%
      dplyr::mutate(mukey = as.character(mukey)),
    by = c("MUKEY" = "mukey")
  ) %>%
  dplyr::mutate(label = paste0(MUKEY, ": ", muname)) %>%
  dplyr::select(-musym) %>%
  plot_map(
    zcol = "label",
    legend = FALSE
  )
```

#### SSURGO by Soil Survey Area
```{r SSURGO-area, message=FALSE}
# Or, download by Soil Survey Area names
SSURGO.areas <- get_ssurgo(
  template = c("CO670"),
  label = "CO670"
)

SSURGO.areas$spatial %>%
  rmapshaper::ms_simplify() %>%
  dplyr::left_join(
    SSURGO.areas$tabular$muaggatt %>%
      dplyr::mutate(mukey = as.character(mukey)),
    by = c("MUKEY" = "mukey")
  ) %>%
  dplyr::mutate(label = paste0(MUKEY, ": ", muname)) %>%
  dplyr::select(-musym) %>%
  plot_map(
    zcol = "brockdepmin",
    layer.name = "Minimum Bedrock Depth (cm)",
    label = "label"
  )
```

### ITRDB
```{r ITRDB, message=FALSE}
# Get the ITRDB records
# Buffer MEVE, because there aren't any chronologies in the Park
ITRDB <-
  get_itrdb(
    template = FedData::meve %>%
      sf::st_buffer(50000),
    label = "meve",
    measurement.type = "Ring Width",
    chronology.type = "Standard"
  )

plot_map(ITRDB$metadata,
  zcol = "SPECIES",
  layer.name = "Species",
  label = "NAME"
) %>%
  leaflet::setView(
    lat = sf::st_centroid(FedData::meve) %>%
      sf::st_coordinates() %>%
      as.numeric() %>%
      magrittr::extract2(2),
    lng = sf::st_centroid(FedData::meve) %>%
      sf::st_coordinates() %>%
      as.numeric() %>%
      magrittr::extract2(1),
    zoom = 9
  )
```

### USGS National Land Cover Dataset
```{r NLCD, message=FALSE}
# Get the NLCD (USA ONLY)
# Returns a raster
NLCD <- get_nlcd(
  template = FedData::meve,
  year = 2019,
  label = "meve"
)

plot(NLCD)

# Also get the impervious or canopy datasets
NLCD_impervious <-
  get_nlcd(
    template = FedData::meve,
    year = 2019,
    label = "meve_impervious",
    dataset = "impervious"
  )

plot(NLCD_impervious)

NLCD_canopy <-
  get_nlcd(
    template = FedData::meve,
    year = 2019,
    label = "meve_canopy",
    dataset = "canopy"
  )

plot(NLCD_canopy)
```

### Annual National Landcover Dataset
```{r NLCD-annual, message=FALSE}
# Get the NLCD (USA ONLY)
# Returns a raster
NLCD_ANNUAL <-
  get_nlcd_annual(
    template = FedData::meve,
    label = "meve",
    year = 2020,
    product =
      c(
        "LndCov",
        "LndChg",
        "LndCnf",
        "FctImp",
        "ImpDsc",
        "SpcChg"
      )
  )

# Returns a data.frame of all downloaded data
NLCD_ANNUAL

plot(NLCD_ANNUAL$rast[[1]])
```


### USDA NASS Cropland Data Layer
```{r NASS-CDL, message=FALSE}
# Get the NASS (USA ONLY)
# Returns a raster
NASS_CDL <- get_nass_cdl(
  template = FedData::meve,
  year = 2016,
  label = "meve"
)

plot(NASS_CDL)
```

### USGS PAD-US
```{r PADUS, message=FALSE, warning=FALSE}
# Get the PAD-US data (USA ONLY)
PADUS <- get_padus(
  template = sf::st_buffer(FedData::meve, 1000),
  label = "meve"
)

plot_map(
  PADUS$Manager_Name,
  layer.name = "Unit Name",
  zcol = "Unit_Nm"
)

# Optionally, specify your unit names
PADUS_meve_ute <- get_padus(
  template = c("Mesa Verde National Park", "Ute Mountain Reservation"),
  label = "meve_ute"
)

plot_map(
  PADUS_meve_ute$Manager_Name,
  layer.name = "Unit Name",
  zcol = "Unit_Nm"
)
```
